---
title: "Activity Monitoring Analysis"
author: "Rakshith Nayak"
date: "September 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

## Activity Monitoring Analysis
This Report analyzes the data collected from a personal activity monitoring device collected at a 5 minute interval through the day during the months of October and November, 2012 from an anonymous individual.

###1.Loading and pre-processing the data 

```{r echo=TRUE}
##Download and unzip the files in current working directory
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity_monitor.zip")
unzip("./activity_monitor.zip")
activity_data<-read.csv("activity.csv",header = T)
```

### 2.Mean total number of steps taken per day?

Histogram of total number of steps taken per day
```{r  echo=TRUE}
library(ggplot2)

activity_daily_summary<-aggregate(activity_data$steps,by=list(activity_data$date),FUN=sum,na.rm=T)
names(activity_daily_summary)[names(activity_daily_summary) == "Group.1"] <- "Date"
names(activity_daily_summary)[names(activity_daily_summary) == "x"] <- "Steps"
ggplot(activity_daily_summary,aes(x=Steps))+geom_histogram(na.rm = T)
```

Mean/Avg No of Steps taken per day is
```{r echo=TRUE}
mean(activity_daily_summary$Steps)
```

Median of Steps taken per day is
```{r echo=TRUE}
median(activity_daily_summary$Steps)
```

###3.Average daily activity pattern
```{r echo=TRUE}
activity_interval_summary<-aggregate(activity_data$steps,by=list(activity_data$interval),FUN="mean",na.rm=T)
names(activity_interval_summary)[names(activity_interval_summary) == "Group.1"] <- "interval"
names(activity_interval_summary)[names(activity_interval_summary) == "x"] <- "steps"

plot(x= activity_interval_summary$interval,y=activity_interval_summary$steps,type = "l",xlab = "Interval",ylab="Avg No of Steps taken daily",main = "Average daily activity")
```

Maximimun steps are taken during the interval
```{r echo=TRUE}
activity_interval_summary[which.max(activity_interval_summary$steps),"interval"]
```


###4.Imputing missing values
Imputing strategy:NA values are imputed by taking an average of the intervals for all days.

```{r echo=TRUE}
library(dplyr)
activity_data_imputed<-activity_data %>% group_by(interval)%>%mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=T),steps))

activity_daily_summary_imputed<-aggregate(activity_data_imputed$steps,by=list(activity_data_imputed$date),FUN=sum,na.rm=T)
names(activity_daily_summary_imputed)[names(activity_daily_summary_imputed) == "Group.1"] <- "Date"
names(activity_daily_summary_imputed)[names(activity_daily_summary_imputed) == "x"] <- "Steps"
ggplot(activity_daily_summary_imputed,aes(x=Steps))+geom_histogram(na.rm = T)
```

Mean/Avg No of Steps taken per day is
```{r echo=TRUE}
mean(activity_daily_summary_imputed$Steps)
```

Median of Steps taken per day is
```{r echo=TRUE}
median(activity_daily_summary_imputed$Steps)
```

###5.Activity patterns between weekdays and weekends
```{r echo=TRUE}
library(lubridate)
activity_data_imputed$Day<-wday(activity_data_imputed$date,label=TRUE)

activity_data_imputed$daytype<- activity_data_imputed$Day
levels(activity_data_imputed$daytype) <- list(
    weekday = c("Mon", "Tues", "Wed", "Thurs", "Fri"),
    weekend = c("Sun", "Sat"))

activity_data_by_day<-activity_data_imputed%>%group_by(daytype,interval)%>%summarise(averageSteps=mean(steps,na.rm=T))
library(lattice)
xyplot(averageSteps~interval|daytype,data = activity_data_by_day,type='l',layout=(c(1,2)),main="Average Daily Activity Pattern by Type of Day", ylab="Average Number of Steps Taken per Interval",xlab = "interval")

```


